Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-01-23 Manuscript/manuscript.Rmd
manuscript updates

Prior to resampling, we performed a series of random forest classification algorithms on the full features space (all combined predictor types), as well as on type-specific feature spaces (overlap counts, overlap predictors, and log2 distances) to evaluate which predictor types were best able to discriminate between genomic bins that either flanked or did not flank TAD boundaries. It was found that, for 10 kb resolution data, models with distance type predictors had higher AUPRC values compared to other predictor types for both GM12878 and K562 cell lines (Figure 4). This result was observed across all other data resolutions as well (Supplementary Figure 2). There was a gradual decline in AUPRC values as resolution increased with models built on 100 kb resolution domain data having the lowest AUPRC curves, regardless of predictor type (Supplementary Table 3).

Recursive feature elimination was used to establish the optimal number of distance-type features to include in downstream analyses. It was determined that, for 10 kb resolution data, near optimal performance was obtained for models with only 32 predictors for each cell line (Figure 5). Results remained consistent across the other data resolutions (Supplementary Figure 4). Therefore, we reduced the feature space for subsequent models to the top 32 predictors, respective to resolution and cell line. The sets of top predictors chosen varied across resolution and cell line due to differences in the parameters provided in the TAD calling algorithm. For the 10 kb resolution data, the most important predictive features that were robust across cell lines were

>>>>

Prior to resampling, we performed a series of random forest classification algorithms on the full features space (all combined predictor types), as well as on type-specific feature spaces (overlap counts, overlap predictors, and log2 distances) to evaluate which predictor types were best able to discriminate between genomic bins that either flanked or did not flank TAD boundaries. It was found that, for 10 kb resolution data, models with distance type predictors had higher AUPRC values compared to other predictor types for both GM12878 and K562 cell lines (Figure 4). This result was observed across all other data resolutions as well (Supplementary Figure 2). There was a gradual decline in AUPRC values as resolution increased with models built on 100 kb resolution domain data having the lowest AUPRC curves, regardless of predictor type (Supplementary Table 2).

Recursive feature elimination was used to establish the optimal number of distance-type features to include in downstream analyses. It was determined that, for 10 kb resolution data, near optimal performance was obtained for models with only 32 predictors for each cell line (Figure 5). Results remained consistent across the other data resolutions (Supplementary Figure 4). Therefore, we reduced the feature space for subsequent models to the top 32 predictors, respective to resolution and cell line. The sets of top predictors chosen varied across resolution and cell line due to differences in the parameters provided in the TAD calling algorithm (Supplementary Table 3). For the 10 kb resolution data, the most important predictive features that were robust across cell lines were a mixture of transcription factor binding sites, chromatin states, and histone modifications including Smc3, Rad21, CTCF, and H3k9 trimethylation (Figure 6). Transcription factor binding sites Smc3, Rad21, and CTCF were also consitently featured as top predictors when comparing cell lines across the the 25 kb and 50 kb resolutions (Supplementary Figure 5). Results when cell lines were compared across the 100 kb resolutions were less consitent with current literature.
